---
title: "Linear Regression"
output: github_document
---

```{r}
library(tidyverse)
library(p8105.datasets)
```


# Load and clean the Airbnb Data

```{r}
data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb |> 
  mutate(stars = review_scores_location / 2) |> 
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood) |> 
  filter(borough != "Staten Island") |> 
  select(price, stars, borough, neighborhood, room_type)

```


Lets fit a Model

```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)

nyc_airbnb = 
  nyc_airbnb |> 
  mutate(
    borough = fct_infreq(borough),
    room_type = fct_infreq(room_type))

fit = lm(price ~ stars + borough, data = nyc_airbnb)
```


Lets look at the fit:
```{r}
summary(fit)

summary(fit)$coeff

```

Dont do those things ^ instead - Tidy up the output!

```{r}
fit %>% 
  broom::glance()
```
- glance gives you a high level look at the model -- comparing lots of models this is helpful 


Tidy up the coefficients: 
```{r}
fit %>% 
  broom::tidy() %>% 
  mutate(term = str_replace(term, "^borough", "Borough: ")) %>% 
  knitr::kable(digits = 3)

```
- broom tidy makes anything into a tibble -- very useful 


# Fit another model

```{r}

  nyc_airbnb %>% 
  mutate(
    borough = fct_infreq(borough),
    room_type = fct_infreq(room_type))

fit = lm(price ~ stars + borough + room_type, data = nyc_airbnb)

fit %>% 
  broom::tidy()
```


## Quick look at Regression Diagostics 
- mostly boils down to get the residuals and examine them to make sure you dont have any overly skewed/non normal distributions in the residuals 

```{r}
nyc_airbnb %>% 
  modelr::add_residuals(fit) %>% 
  ggplot(aes(x = resid)) +
  geom_density() +
  xlim(-100, 500)

nyc_airbnb %>% 
  modelr::add_residuals(fit) %>% 
  ggplot(aes(x = borough, y = resid)) +
  geom_violin() 


nyc_airbnb %>% 
  modelr::add_residuals(fit) %>% 
  ggplot(aes(x = stars, y = resid)) +
  geom_point() 
```
** you can fit a model on one dataset and then use the model on another dataset to see how it works 

## Hypothesis test for a categorical predictor 

lets fit a Null and Alternative Hypothesis model

```{r}
fit_null = lm(price ~ stars + borough, data = nyc_airbnb)
fit_alt = lm(price ~ stars + borough + room_type, data = nyc_airbnb)

anova(fit_null, fit_alt) %>% 
  broom::tidy()
```


## Borough Level Difference

```{r}

```

